---
title: "Results: Algorithm 1"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding) { 
  out_dir <- '../../docs/';
  rmarkdown::render(inputFile,
                    encoding=encoding, 
                    output_file=file.path(dirname(inputFile), out_dir, 'algorithm1.html')) })
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
here::i_am("code/instances/algorithm1.Rmd")  # specify relative path given project
# remotes::install_github("relund/gMOIP")
# remotes::install_github("fischuu/OjaNP")
# library(gMOIP)
library(DT)
library(tidyverse)
# library(knitr)
# library(rgl)
# rgl::setupKnitr()
# options(rgl.useNULL=TRUE)
# rgl::par3d("family" = "serif")
knitr::opts_chunk$set(
  echo = FALSE,
  message=FALSE, include = TRUE, warning = FALSE,
  out.width = "99%", fig.width = 12, fig.align = "center", fig.asp = 0.8
)
```

This report contains different plots and tables that may be relevant for analysing the results. Observe:

  * The hypercube $Y_N$ is placed in, is approx. $[0, 10000m]^p$, since for each subproblem $Y_N^s \subset [0, 10000]^p$. That is, more subproblems create a larger hypercube for the master problem and the volume (number of integer points) grows exponentially in $p$. 
  * Given $|Y_N|$, we can find the hypercube as $$[l_1, u_1] \times \cdots \times [l_p, u_p],$$ where $l_i = \min_{y \in Y_N}(y_i)$ and $u_i = \max_{y \in Y_N}(y_i)$.
  * The number of integer points in the hypercube is $$\prod_{i = 1}^p (u_i - l_i + 1)$$


## Statistics for `alg1`

Given a problem consisting of $m$ subproblems with $Y_N^s$ given for each subproblem $s$, we use a filtering algorithm to find $Y_N$ (`alg1`). 

The following instance/problem groups are generated given:

  * $p=2,\ldots, 5$. [4 options]
  * $m=2,\ldots 5$ where $m$ is the number of subproblems. [4 options]
  * All subproblems have the same method config or half have method `u` and `l`. [4 options]
  * Five instances for each config. [5 options]

### Status 

```{r}
files <- fs::path_file(fs::dir_ls(here::here("code/instances/problems")))
files <- str_c("alg1-", files)
raw <- read_csv(here::here("code/instances/stat-prob.csv")) |> mutate(path = fs::path_file(path))
raw <- full_join(tibble(path = files), raw)
dat <- raw %>% 
  mutate(method = str_extract(path, "^.*-(.*)-.*$", group = 1)) %>% 
  mutate(method = map_chr(method, function (x) {
  x <- str_split_1(x, "")
  x <- unique(x)
  str_c(x, collapse = "")
  })) %>% 
  rowwise() %>% 
  mutate(aveWidth = mean(c_across(starts_with("width")), na.rm = T),
         points_hypercube = if_else(is.nan(aveWidth), NA, prod(1 + c_across(starts_with("width")), na.rm = T)),
         spAveCard = mean(c_across(starts_with("spCard")), na.rm = T)) %>% 
  ungroup() %>% 
  mutate(
         pctSup = supported/card,
         pctExt = extreme/card,
         m = as.numeric(str_extract(path, "^.*-(.*)_.*$", group = 1))) 


## is the hypercube close to approx?
#dat %>% mutate(cst = points_hypercube/(m * 10000)^p) %>% ungroup() %>% summarise(val = mean(cst, na.rm = T))

# add relative stat for subproblems
for (s in 1:max(dat$m)) {
  dat[, str_c("spPctExt", s)] <- dat[, str_c("spExtreme", s)] / dat[, str_c("spCard", s)]
} 
dat <- dat %>% 
  rowwise() %>% 
  mutate(
    spAvePctExt = mean(c_across(starts_with("spPctExt")), na.rm = T),
    ubCard = prod(c_across(starts_with("spCard")), na.rm = T)
  ) %>% 
  ungroup()
datFull <- dat
# remove results where have not solved all instances
dat <- dat %>% 
  group_by(p, m, method, spAveCard) %>% 
  mutate(instances = n()) %>% 
  filter(instances == 5, spAveCard <= 600) %>% 
  select(-instances) %>% 
  ungroup()
```

  * `r datFull |> filter(!is.na(card)) |> nrow()`/`r nrow(datFull)` problems have been solved, i.e. `r nrow(datFull) - datFull |> filter(!is.na(card)) |> nrow()` remaining.
  * `r nrow(dat)`/`r datFull |> filter(!is.na(card)) |> nrow()` problems have 5 instances solved for each configuration.
  * `r datFull |> filter(is.na(extreme) & !is.na(card)) |> nrow()`/`r datFull |> filter(!is.na(card)) |> nrow()` have not been classified.
  * `r datFull |> filter(is.na(unsupported) & !is.na(extreme)) |> nrow()`/`r datFull |> filter(!is.na(card)) |> nrow()` have not been fully classified (only classified extreme).
  
<!-- Overview over subproblems not solved yet -->

<!-- ```{r} -->
<!-- dat %>% -->
<!--   filter(is.na(card)) |>  -->
<!--   group_by(p, m, method, spAveCard) %>% -->
<!--   count() -->
<!-- ``` -->
  
### Problems solved for the analysis  

```{r, out.width = "100%"}
DT::datatable(dat, width = "100%", options = list(autoWidth = T, scrollX = TRUE, columnDefs = list(list(width = '250px', targets = c(1)))))
```



```{r}

time_limit <- 1*30*60

my_dat <- dat %>% 
   mutate(under_limit = (filter_time < time_limit)) %>%
   filter(filter_time < 40000)


my_dat %>% filter(under_limit == FALSE) %>% select(path)

my_dat %>%    group_by(under_limit) %>%
 summarise(
    true_count = sum(under_limit == TRUE),
    false_count = sum(under_limit == FALSE)
   )

my_dat$under_limit <- as.factor(my_dat$under_limit)

my_dat %>% select(path, filter_time, under_limit) %>% filter(under_limit == FALSE) %>% summarise(max = max(filter_time))

ggplot(my_dat, aes(x = filter_time), color=under_limit) +
  geom_histogram(binwidth = 5, fill = "blue") +
  labs(title = "Histogram of filter times", x = "filter time", y = "Frequency", colors = 'Under_limit')

```



Note that the width of objective $i = 1, \ldots p$, $w_i = [l_i, u_i]$ should be approx. $10000m$. Check:

```{r}
dat %>% 
  group_by(m) %>% 
  summarize(across(starts_with("width"), ~ mean(.x, na.rm = TRUE), .names = "mean_{.col}"))
```



```{r, eval=FALSE}
# pts <- matrix(c(5,5,5), ncol = 3)
# ini3D(argsPlot3d = list(xlim = c(0,10), ylim = c(0,10), zlim = c(0,10)))
# plotCones3D(pts)
# plotCones3D(pts, direction = -1)
# finalize3D(argsAxes3d = list(edges = "bbox"))

## try to estimate ub on integer nd points in a hypercube by fitting pts to hyperplane

# library(eaf)
# library(OjaNP)
# estimateUbND <- function(ideal, nadir) {
#   p <- length(ideal)
#   # find lex max
#   pts <- matrix(rep(ideal, p), ncol = p, byrow = T)
#   diag(pts) <- nadir
#   # calc hyperplane containing pts
#   coeff <- hyperplane(pts)   # d^t * x + c = 0
#   # find integer pts near hyperplane
#   val <- map(1:(p-1), function(i) ideal[i]:nadir[i])
#   res <- as.matrix(expand.grid(val))
#   coeff1 <- coeff[1:(p-1)]
#   tmp <- -(coeff1 %*% t(res) + coeff[p+1])/coeff[p]
#   res <- cbind(res,tmp[1,])
#   # tst <- coeff[1:p] %*% t(tst) + coeff[p]
#   # any(tst != 0)
#   res <- res[res[,p] >= ideal[p] & res[,p] <= nadir[p], ]
#   # colnames(res) <- str_c("y", 1:(p-1))
#   # res <- res |> 
#   #   rowwise() |> 
#   #   mutate(test = -(sum(coeff[1:(p-1)] * c_across(contains("y"))) + coeff[p+1]) / coeff[p]) |> 
#   #   mutate(y = round(-(sum(coeff[1:(p-1)] * c_across(contains("y"))) + coeff[p+1]) / coeff[p])) |> 
#   #   mutate(test = sum(coeff[1:p] * c_across(contains("y"))) + coeff[p+1]) 
#   ndSet <- filter_dominated(res)
#   return(nrow(ndSet))
# }
# p <- 3
# ideal <- rep(0,p)
# nadir <- rep(10000,p)
# estimateUbND(ideal, nadir)
```



### Size of $Y_N$

What is $|Y_N|$ given the different methods of generating the set of nondominated points for the subproblems?

```{r}
dat %>% 
  group_by(method) %>% 
  summarise(mean_card = mean(card), n = n()) 
```

Does $p$ have an effect? 

```{r}
dat %>% 
  group_by(method, p) %>% 
  summarise(mean_card = mean(card), n = n()) %>% 
  arrange(p)
```

Does $m$ have an effect? 

```{r}
dat %>% 
  group_by(method, m) %>% 
  summarise(mean_card = mean(card), n = n()) %>% 
  arrange(m)
```

```{r}
dat %>% 
  ggplot(aes(x = spAveCard, y = card, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, scales = "free_y", labeller =  "label_both")  + 
  labs(title = "Number of nondominated points given subproblem cardinality",
       y = "Cardinality Yn",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") 
```

```{r}
dat %>% 
  ggplot(aes(x = spAveCard, y = card, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, scales = "free_y", labeller = "label_both")  + 
  labs(title = "Number of nondominated points given subproblem cardinality (log2) scale",
       y = "Cardinality Yn",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") +
  scale_y_continuous(trans = "log2")
```

<!-- ```{r} -->
<!-- dat %>%  -->
<!--   ggplot(aes(x = spAveCard, y = card, color = factor(p))) + -->
<!--   geom_point() +  -->
<!--   stat_summary(fun="mean", geom="line") + -->
<!--   facet_wrap(vars(method), scales = "free", labeller = "label_both")  +  -->
<!--   labs(title = "Number of nondominated points given subproblem cardinality", -->
<!--        y = "Cardinality Yn", -->
<!--        x = "Subproblem cardinality", -->
<!--        color = "p") + -->
<!--   theme(legend.position = "bottom")  -->
<!-- ``` -->

```{r}
dat %>% 
  ggplot(aes(x = spAvePctExt, y = card, color = factor(spAveCard), shape = method, group = method)) +
    stat_summary(fun="mean", geom="line", color = "gray70") +
    # stat_summary(aes(group = factor(spAveCard), color = factor(spAveCard)), fun="mean", geom="line", size = 0.05) +
    geom_point() + 
    facet_grid(p ~ m, scales = "free_y", labeller =  "label_both")  + 
    labs(title = "Number of nondominated points",
       y = "Cardinality Yn",
       x = "Ave relative number of subproblem extreme",
       color = "Subproblem cardinality") +
  theme(legend.position = "bottom") 
```


### Relative size of $Y_N$

```{r}
dat %>% 
  ggplot(aes(x = spAveCard, y = card / ubCard, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, labeller = "label_both")  + 
  labs(title = "Relative size of nondominated points compared to total enumeration given subproblem cardinality",
       y = "Cardinality Yn",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") 
```

```{r, eval=FALSE}
dat %>% 
  ggplot(aes(x = spAveCard, y = card / ubCard, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, labeller = "label_both")  + 
  labs(title = "Relative size of nondominated points compared to total enumeration given subproblem cardinality (log2 scale)",
       y = "Cardinality Yn",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") +
  scale_y_continuous(trans = "log2")
```

```{r}
dat %>% 
  ggplot(aes(y = card / ubCard, x = method, color = method)) +
  geom_boxplot() +
  facet_grid(p ~ m, labeller = label_both) +
  labs(title = "Relative size of nondominated points compared to total enumeration",
       y = "%",
       x = "Method",
       color = "Method(s) used to generate subproblems") +
  theme(legend.position = "bottom")
```


```{r, eval=FALSE}
dat %>% 
  ggplot(aes(y = card / ubCard, x = method, color = method)) +
  geom_boxplot() +
  facet_grid(m ~ p, labeller = label_both) +
  labs(title = "Relative size of nondominated points compared to total enumeration (log2 scale)",
       y = "%",
       x = "Method",
       color = "Method(s) used to generate subproblems") +
  theme(legend.position = "bottom") +
  scale_y_continuous(trans = "log2")
```


<!-- ```{r} -->
<!-- dat %>%  -->
<!--   ggplot(aes(y = card / points_hypercube, x = method, color = method)) + -->
<!--   geom_boxplot() + -->
<!--   facet_grid(m ~ p, margins = T, labeller = label_both) + -->
<!--   labs(title = "Relative size of nondominated points compared to total enumeration", -->
<!--        y = "%", -->
<!--        x = "Method", -->
<!--        color = "Method(s) used to generate subproblems") + -->
<!--   theme(legend.position = "bottom") -->
<!-- ``` -->


### Nondominated points classification

We classify the nondominated points into, extreme, supported non-extreme and unsupported.

```{r}
dat %>% 
  ggplot(aes(x = spAveCard, y = pctExt, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, margins = T, labeller = label_both)  + 
  labs(title = "Relativ number of extreme points given subproblem cardinality",
       y = "%",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") 
```

```{r}
dat %>% 
  ggplot(aes(x = spAveCard, y = pctSup, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, margins = T, labeller = label_both)  + 
  labs(title = "Relativ number of supported points given subproblem cardinality",
       y = "%",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") 
```

```{r}
dat %>% 
  ggplot(aes(x = method, y = pctExt, color = method)) +
  geom_boxplot() + 
  facet_grid(p ~ m, margins = T, labeller = label_both) +
  labs(title = "Relativ number of extreme points given method",
       y = "%",
       x = "Method",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom")
```

```{r}
dat %>% 
  ggplot(aes(x = spAvePctExt, y = pctExt, color = method)) +
  geom_point() + 
  facet_grid(p ~ m, margins = T, labeller = label_both) +
  labs(title = "Relativ number of extreme points",
       y = "% extreme master",
       x = "Avg. % extreme subproblems",
       color = "Method(s) used to generate subproblems") +
  theme(legend.position = "bottom")
```


```{r}
dat %>% 
  ggplot(aes(x = spAvePctExt, y = pctExt, color = factor(spAveCard), shape = method)) +
  geom_point() + 
  facet_grid(p ~ m, margins = T, labeller = label_both) +
  labs(title = "Relativ number of extreme points",
       y = "% extreme master",
       x = "Avg. % extreme subproblems",
       color = "Subproblem cardinality") +
  theme(legend.position = "bottom")
```


```{r}
dat %>% 
  ggplot(aes(y = pctExt/spAvePctExt, x = spAveCard, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, margins = T, labeller = label_both) +
  labs(title = "Relativ: % extreme master / % extreme subproblems",
       y = "% extreme master / % extreme subproblems",
       x = "Subproblem cardinality",
       color = "Method(s) used to generate subproblems") +
  theme(legend.position = "bottom")
```



## Plots used in the paper

```{r, eval=FALSE}
## Number of extreme

## how many sne?
dat |> 
  filter(card != 10, card != 150) |> 
  mutate(e = extreme/card, sne = (supported-extreme)/card) |> 
  group_by(method, p) |> 
  summarise(e = mean(e), sne = mean(sne))

dat1 <- dat |> 
  filter(card != 10, card != 150) 

pt1 <- dat1 |> 
  ggplot(aes(x = card, y = extreme/card, shape = method, linetype = method)) +
  stat_summary(fun="mean", geom="line") + 
  geom_point(size = 0.75) + 
  facet_grid(cols = vars(p)) +
  labs(x = "$\\Yn^s$", y = NULL) +
  theme_bw() +
  theme(legend.position = "none")
pt1

pt2 <- dat1 |> 
  ggplot(aes(x = p, y = extreme/card, shape = method, linetype = method)) +
  geom_point(size = 0.75) + 
  stat_summary(fun="mean", geom="line") + 
  facet_grid(cols = vars(card)) + 
  labs(x = "$p$", y = NULL) +
  theme_bw() +
  theme(legend.position = "none")
pt2

tikz("se-card.tex", width = 8, height = 3)
pt1
dev.off()
tikz("se-p.tex", width = 8, height = 3)
pt2
dev.off()
```
