---
title: "Generation of instance files for subproblems"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
knit: (function(inputFile, encoding) { 
    out_dir <- '../../docs/';
    rmarkdown::render(inputFile,
                      encoding=encoding, 
                      output_file=file.path(dirname(inputFile), out_dir, 'problem-generation.html')) })
---

```{r setup, include=FALSE}
remotes::install_github("relund/gMOIP")
library(gMOIP)
library(tidyverse)
library(knitr)
library(rgl)
rgl::setupKnitr()
options(rgl.useNULL=TRUE)
rgl::par3d("family" = "serif")
opts_chunk$set(
  echo = FALSE,
  message=FALSE, include = TRUE,
  out.width = "99%", fig.width = 8, fig.align = "center", fig.asp = 0.8
)
```

## Generation of subproblems

```{r}
genSubproblems <- function(p, samples, width = 10000, pltS = FALSE, pltM = FALSE) {
  #### subproblems sphere ####
  center <- rep(0, p)
  radius <- width
  pts <- genSample(p, samples * 10 * p, 
                   argsSphere = list(center = center, radius = radius, below = NULL, factor = 6))
  # cat("Genenerate", dim(pts)[1], "on a sphere\n")
  ## Upper part
  ptsU <- as.data.frame(pts) %>%
    filter(if_all(where(is.double), ~ (. >= 0))) %>% 
    slice_sample(n = round(1.2 * samples))
  # cat("Pick ND points among", dim(ptsU)[1], "upper points\n")
  ptsU <- addNDSet(ptsU[, 1:p], crit = "min", classify = F)
  if (nrow(ptsU) < samples) warning("Did not find ", samples, " nd points (u). p = ", p, " s = ", samples)
  ptsU <- ptsU %>% 
    slice_sample(n = samples)
  # cat("Classify", dim(ptsU)[1], "upper points\n")
  ptsU <- classifyNDSet(ptsU[,1:p]) %>% mutate(nd = TRUE)
  ## lower part
  ptsL <- as.data.frame(pts) %>%
    filter(if_all(where(is.double), ~ (. <= 0))) %>% 
    slice_sample(n = round(1.2 * samples))
  # cat("Pick ND points among", dim(ptsL)[1], "lower points\n")
  ptsL <- addNDSet(ptsL[, 1:p], crit = "min", classify = F)
  if (nrow(ptsL) < samples) warning("Did not find ", samples, " nd points (l). p = ", p, " s = ", samples)
  ptsL <- ptsL %>% 
    slice_sample(n = samples)
  # cat("Classify", dim(ptsL)[1], "lower points\n")
  ptsL <- classifyNDSet(ptsL[,1:p]) %>% mutate(nd = TRUE)
  ## Plot
  if (p == 3 & pltS) {
    # rgl::mfrow3d(nr = 1, nc = 2, sharedMouse = F)
    ini3D(argsPlot3d = list(box = TRUE, axes = TRUE))
    rgl::spheres3d(center, radius=radius, color = "grey100", alpha=0.1)
    plotPoints3D(ptsU)
    plotPoints3D(ptsL)
    finalize3D()
    # next3d()
  }
  ptsL <- ptsL %>% 
    mutate(across(where(is.double), ~ .x + width)) 
  Rfast::colMinsMaxs(as.matrix(ptsU[, 1:p]))
  Rfast::colMinsMaxs(as.matrix(ptsL[, 1:p]))
  
  #### subproblems planes ####
  rng <- matrix(rep(c(0, width), each = p), nrow = p)
  center <- rowMeans(rng)
  planeU <- c(rep(1, p), -1.2*sum(rowMeans(rng)))
  planeL <- c(rep(1, p), -0.8*sum(rowMeans(rng)))
  ptsM <- genNDSet(p, n = samples, range = rng, planes = TRUE, classify = TRUE,
                   argsPlanes = list(center = center, planeU = planeU, planeL = planeL))
  # ptsR <- genSample(p, 10 * p * samples, range = rng, random = TRUE, classify = T)
  # center <- rep(width/2, p)
  # planeL <- c(center+width/4, -sum(center^2))
  # planeU <- c(center-width/15, -sum(center^2))
  # set <- matrix(rep(0, p), nrow = 1)
  # for (i in 1:nrow(ptsR)) {
  #   if (sum(planeL[1:p] * ptsR[i, ]) >= 
  #     -planeL[p + 1]) {
  #     set <- rbind(set, ptsR[i, ])
  #   }
  # }
  # ptsR <- set[-1, ]
  # set <- matrix(rep(0, p), nrow = 1)
  # for (i in 1:nrow(ptsR)) {
  #   if (sum(planeU[1:p] * ptsR[i, ]) <= 
  #     -planeU[p + 1]) {
  #     set <- rbind(set, ptsR[i, ])
  #   }
  # }
  # ptsR <- set[-1, ]
  # ptsR <- as.data.frame(ptsR) %>%
  #   slice_sample(n = round(4 * p * samples))
  # # cat("Pick ND points among", dim(ptsR)[1], "middle points\n")
  # ptsR <- addNDSet(ptsR, crit = "min", classify = F)
  # if (nrow(ptsR) < samples) warning("Did not find ", samples, " nd points (m). p = ", p, " s = ", samples)
  # ptsR <- ptsR %>% 
  #   slice_sample(n = samples)
  # # cat("Classify", dim(ptsR)[1], "middle points\n")
  # ptsR <- classifyNDSet(ptsR[,1:p])
  # Rfast::colMinsMaxs(as.matrix(ptsR[, 1:p]))
  if (p == 3 & pltM) {
    ini3D(argsPlot3d = list(box = TRUE, axes = TRUE))
    plotPoints3D(ptsM)
    rgl::planes3d(planeL[1],planeL[2],planeL[3],planeL[4], alpha = 0.5, col = "red")
    rgl::planes3d(planeU[1],planeU[2],planeU[3],planeU[4], alpha = 0.5, col = "red")
    finalize3D()
  }
  
  
  #### subproblems box ####
  # rng <- matrix(rep(c(0, width, 0, round(width/4, 0)), p)[1:(2 * p)], byrow = TRUE, ncol = 2)
  # pts <- genSample(p, samples * 2, range = rng, box = TRUE, argsBox = list(cor = "idxRand"))
  # ptsB <- pts
  # if (p == 3 & plt) {
  #   ini3D(argsPlot3d = list(box = TRUE, axes = TRUE))
  #   plotPoints3D(ptsB)
  #   finalize3D()
  # }
  
  calcStat <- function(p, pts, method) {
    lst <- list(points = pts %>% select(-nd, -se, -sne, -us))
    lst$statistics$p <- p 
    lst$statistics$card <- nrow(pts)
    lst$statistics$supported <- sum(pts$se) + sum(pts$sne)
    lst$statistics$extreme <- sum(pts$se)
    lst$statistics$unsupported <- sum(pts$us)
    lst$statistics$min <- Rfast::colMins(as.matrix(pts[, 1:p]), value = T)
    lst$statistics$max <- Rfast::colMaxs(as.matrix(pts[, 1:p]), value = T)
    lst$statistics$width <- Rfast::colrange(as.matrix(pts[, 1:p]))
    lst$statistics$method <- method
    return(lst)
  }
  
  return(list(iU = calcStat(p, ptsU, "u"), iM = calcStat(p, ptsM, "m"), iL = calcStat(p, ptsL, "l")))
}

saveSubproblems <- function(pS, samplesS, dub, overwrite = TRUE) {
  #name = sub-<p>-<card>-<gen-method>_-<grp-id>.json
  for (p in pS) {
    for (s in samplesS) {
      for (d in 1:dub) {
        pathU <- str_c("subproblems/sp-", p, "-", s, "-u", "_", d, ".json")
        pathM <- str_c("subproblems/sp-", p, "-", s, "-m", "_", d, ".json")
        pathL <- str_c("subproblems/sp-", p, "-", s, "-l", "_", d, ".json")
        if (overwrite | !fs::file_exists(pathU) | !fs::file_exists(pathM) | !fs::file_exists(pathL)) {
          cat("Generate for p =", p, "card =", s, "\n")
          lst <- genSubproblems(p, s)
          jsonlite::write_json(lst$iU, pathU, pretty = TRUE)
          jsonlite::write_json(lst$iM, pathM, pretty = TRUE)
          jsonlite::write_json(lst$iL, pathL, pretty = TRUE)
        }
      }
    }
  }
  return(NULL)
}

getStatistics <- function(paths = NULL) {
  dat <- NULL
  if (is.null(paths)) {
    paths <- fs::dir_ls("subproblems/")
    paths <- paths[str_detect(paths, "-")]
    pahts <- sort(paths, decreasing = T)
  }
  for (p in paths) {
    lst <- jsonlite::read_json(p)
    dat <- dat %>% bind_rows(c(path = p, unlist(lst$statistics)))
  }
  return(dat)
}

deleteFilesProb <- function(delete = TRUE) {
  paths <- fs::dir_ls("problems", glob = "*/prob*")
  paths <- sort(paths, decreasing = T)
  grps <- unique(str_extract(paths, "^(.*)_.*.json$", group = 1))
  for (g in grps) {
    paths1 <- str_subset(paths, fixed(g))
    checksums <- 0
    for (p in paths1) {
      subp <- sort(jsonlite::read_json(p, simplifyVector = T))
      csum <- sum(rep(1:length(subp)) * as.numeric(str_extract(subp, "^.*_(.*).json$", group = 1)))
      if (csum %in% checksums) {
        cat("Remove dublicate", p, "!\n")
        fs::file_delete(p)
      }
      checksums <- c(checksums, csum)
    }
  }
  return(NULL)
} 
  
  
deleteFiles <- function(delete = TRUE) {
  paths <- fs::dir_ls("subproblems/")
  paths <- str_subset(paths, "sp-")  # remove statistics file
  if (delete == "m") {
     fs::file_delete(str_subset(paths, "-m"))
  }
  for (p in sort(paths, decreasing = T)) {
    samples <- as.numeric(str_split_1(p, "-")[3])
    lst <- jsonlite::read_json(p)
    if (samples != lst$statistics$card[[1]]) {
      cat("Found lower number of nd points for", p, samples, "!=", lst$statistics$card[[1]],  " Remove the file!\n")
      if (delete) fs::file_delete(p)
    }
  }
  return(NULL)
} 
```

We use the R package `gMOIP` to generate subproblems. A subproblem is generated such that all nondominated points is integer and in the hypercube $[0, 10000]^p$. A configuration is defined using:

  * Number of objectives ($p$).
  * Number of nondominated points (card). 
  * Generation method which is either
    - Points generated on the upper (u) part of a sphere resulting in many unsupported points (see Fig. 1).
    - Points generated between to hyperplanes in the middle (m) of the hypercube, resulting in both supported and unsupported points near to the hull (see Fig. 2).
    - Points generated on the lower (l) part of a sphere resulting in many supported points (see Fig. 1).
  
That is a subproblem can be identified using filename `sp-<p>-<card>-<gen-method>_<id>.json` where `id` denote the instance id for the same configuration. 
  
```{r, webgl=TRUE}
lst <- genSubproblems(p = 3, samples = 25, pltS = T)
rgl::title3d(main = "Fig. 1: Generation on a sphere (option u and l).")
```

```{r, webgl=TRUE}
lst <- genSubproblems(p = 3, samples = 25, pltM = T)
rgl::title3d(main = "Fig. 2: Generation between hyperplanes (option m).")
```
  
For each subproblem we calculate the statistics:

  * Classification of points into supported extreme (se), supported non extreme (sne), unsupported (us).
  * Number of 
    - points (cardinality) 
    - supported, extreme and unsupported 
    <!-- - ratio $r$ = unsupported/supported. Note unsupported = $r\cdot$supported, $r\in [0, unsupported]$. If $r$ = 0 (no unsupported), $r=$unsupported (one supported), $r<1$ more supported.  -->
    - min and max value for each objective $i = 1,\ldots p$.
    - width $w_i$ = $max_i-min_i$ for each objective $i = 1,\ldots p$.

Let us try to generate a set of subproblems saved as json files in folder subproblems:

```{r, echo=TRUE, results='hide'}
pS <- 2:5 # p values
samplesS <- c(10, 50, 100, 150, 200, 300, 600) # nd points set size
dub <- 10 # number of duplicates for each configuration
deleteFiles()
saveSubproblems(pS, samplesS, dub, overwrite = F)
```

```{r, include=FALSE}
path <- "subproblems/statistics.csv"
dat <- getStatistics()
if (!is.null(dat)) write_csv(dat, file = path) 
dat <- read_csv(path)
```

### A few plots

In total there are `r nrow(dat)` subproblems. 

```{r}
ggplot(mapping = aes(x = card, y = unsupported/card, color = method), data = dat) +
  geom_point() + 
  geom_smooth(se = F) +
  facet_wrap(vars(p)) + 
  labs(title = "Percentage of unsupported given p",
       y = "Pct.") +
  theme(legend.position = "bottom")

dat %>% 
  rowwise() %>% 
  mutate(avg_width = max(c_across(contains("width")), na.rm = TRUE)) %>% 
  ggplot(mapping = aes(x = card, y = avg_width, color = method)) +
  geom_point() + 
  geom_smooth(se = F) +
  facet_wrap(vars(p)) + 
  labs(title = "Average width for all objectives given p") +
  theme(legend.position = "bottom")
```


## Generation of problems

The problems generated consists of $m\leq 5$ subproblems and is generated so provide a good test bed for a research question. The naming convention is `prob-<p>-<cards>-<gen-methods>_<m>_<id>.json`

### RQ: How large is $|Y_N|$ given $|y_N^s|, s\in S$?

The following instance/problem groups are generated given:

  * $p=2,\ldots, 5$. [4 options]
  * $m=2,\ldots 5$ where $m$ is the number of subproblems. [4 options]
  * All subproblems have the same method config or half have method `u` and `l`. [4 options]
  * Five instances for each config. [5 options]

```{r}
saveProblemsAlg1 <- function(pS, mS, samplesS, methodS, dub, overwrite = TRUE) {
  paths <- fs::dir_ls("subproblems/")
  paths <- str_subset(paths, "sp-")  # remove redundant files
  for (p in pS) {
    for (s in samplesS) {
      for (method in methodS) {
        if (method != "ul") {
          subS <- str_c("subproblems/sp-", p, "-", s, "-", method)
          curPaths <- str_subset(paths, subS)
        } 
        if (method == "ul") {
          subS <- str_c("subproblems/sp-", p, "-", s, "-u")
          curPathsU <- str_subset(paths, subS)
          subS <- str_c("subproblems/sp-", p, "-", s, "-l")
          curPathsL <- str_subset(paths, subS)
        }
        for (m in mS) {
          datSubp <- as.data.frame(matrix(NA_character_, nrow = 10 * dub, ncol = m), .name_repair = "minimal")
          for (d in 1:(10 * dub)) {
            if (method != "ul") datSubp[d, ] <- sample(curPaths, m)
            if (method == "ul") {
              mL <- round(m/2)
              mU <- m - mL
              subp <- sample(curPathsU, mU)
              subp <- c(subp, sample(curPathsL, mL))
              datSubp[d, ] <- subp
            }
          }
          datSubp <- datSubp %>% distinct() %>% slice_sample(n = dub)
          for (d in 1:dub) {
            subp <- datSubp[d, ]
            dat <- getStatistics(subp)
            ss <- str_c(dat$card, collapse = "|")
            mth <- str_c(dat$method, collapse = "")
            path <- str_c("problems/prob-", p, "-", ss, "-", mth, "-", m, "_", d, ".json")
            if (overwrite | !fs::file_exists(path) ) {
              cat("Save", path, "\n")
              jsonlite::write_json(subp, path, pretty = TRUE)
            }
          }
        }
      }
    }
  }
  return(NULL)
}
```


```{r, echo=TRUE, results='hide'}
pS <- 2:5
mS <- 2:5
samplesS <- c(50, 200, 300, 600) # nd points set size
methodS <- c("u", "l", "m", "ul")
dub <- 5 # number of duplicates for each configuration
saveProblems(pS, mS, samplesS, methodS, dub, overwrite = F)
```






