---
title: "Results: Plots and Tables"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
# knit: (function(inputFile, encoding) { 
#     out_dir <- '../../docs/';
#     rmarkdown::render(inputFile,
#                       encoding=encoding, 
#                       output_file=file.path(dirname(inputFile), out_dir, 'problem-statistics.html')) })
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup, include=FALSE}
here::i_am("code/instances/problem-statistics.Rmd")  # specify relative path given project
# remotes::install_github("relund/gMOIP")
# remotes::install_github("fischuu/OjaNP")
# library(gMOIP)
library(DT)
library(tidyverse)
# library(knitr)
# library(rgl)
# rgl::setupKnitr()
# options(rgl.useNULL=TRUE)
# rgl::par3d("family" = "serif")
conflicted::conflicts_prefer(plotly::layout)
conflicted::conflicts_prefer(dplyr::filter)
knitr::opts_chunk$set(
  echo = FALSE,
  message=FALSE, include = TRUE, warning = FALSE,
  out.width = "99%", fig.width = 12, fig.align = "center", fig.asp = 0.8
)

library(tikzDevice)
options(tikzLatexPackages = c(
   getOption('tikzLatexPackages'),
   "\\usepackage{xspace}",
   "\\newcommand{\\Y}{\\mathcal{Y}}",
   "\\newcommand{\\Yn}{\\mathcal{Y}_{\\textnormal{N}}}",
   "\\newcommand{\\Yns}{\\mathcal{Y}_{\\textnormal{s}}}",
   "\\newcommand{\\Ynse}{\\mathcal{Y}_{\\textnormal{se}}}",
   "\\newcommand{\\Ynsne}{\\mathcal{Y}_{\\textnormal{sne}}}",
   "\\newcommand{\\Ynu}{\\mathcal{Y}_{\\textnormal{u}}}",
   "\\newcommand{\\Yd}{\\mathcal{Y}_{\\textnormal{d}}}",
   "\\newcommand{\\mL}{\\texttt{\\textsc{l}}\\xspace}",
   "\\newcommand{\\mU}{\\texttt{\\textsc{u}}\\xspace}",
   "\\newcommand{\\mM}{\\texttt{\\textsc{m}}\\xspace}"
))
```

This report contains different plots and tables that may be relevant for analysing the results. Observe:

  * The hypercube $Y_N$ is placed in, is approx. $[0, 10000m]^p$, since for each subproblem $Y_N^s \subset [0, 10000]^p$. That is, more subproblems create a larger hypercube for the master problem and the volume (number of integer points) grows exponentially in $p$. 
  * Given $|Y_N|$, we can find the hypercube as $$[l_1, u_1] \times \cdots \times [l_p, u_p],$$ where $l_i = \min_{y \in Y_N}(y_i)$ and $u_i = \max_{y \in Y_N}(y_i)$.
  * The number of integer points in the hypercube is $$\prod_{i = 1}^p (u_i - l_i + 1)$$


## Statistics for `alg1`

Given a problem consisting of $m$ subproblems with $Y_N^s$ given for each subproblem $s$, we use a filtering algorithm to find $Y_N$ (`alg1`). 

The following instance/problem groups are generated given:

  * $p=2,\ldots, 5$. [4 options]
  * $m=2,\ldots 5$ where $m$ is the number of subproblems. [4 options]
  * All subproblems have the same method config or half have method `u` and `l`. [4 options]
  * Five instances for each config. [5 options]

### Status 

```{r}
extractInfo <- function(path) {
  path <- fs::path_file(path)
  dat <- tibble(path = path) |> 
    mutate(
      p = as.numeric(str_extract(path, "^alg1-prob-(.).*$", group=1)),
      m = as.numeric(str_extract(path, "^.*(.)_..json$", group=1)),
      spAveCard = as.numeric(str_extract(path, "^.*-(.*?)\\|.*$", group = 1)),
      method = str_extract(path, "^.*-(.*)-.*$", group = 1) %>% 
        map_chr(function (x) {
          x <- str_split_1(x, "")
          x <- unique(x)
          str_c(x, collapse = "")
        })
    )
  return(dat)
}

files <- fs::path_file(fs::dir_ls(here::here("code/instances/problems")))
files <- str_c("alg1-", files)
files <- extractInfo(files)
raw <- read_csv(here::here("code/instances/stat-prob.csv")) |> 
  mutate(path = fs::path_file(path)) |> 
  distinct(path, .keep_all = T)
raw <- full_join(files, raw)
dat <- raw %>% 
  mutate(method = str_extract(path, "^.*-(.*)-.*$", group = 1)) %>% 
  mutate(method = map_chr(method, function (x) {
    x <- str_split_1(x, "")
    x <- unique(x)
    str_c(x, collapse = "")
  })) %>% 
  mutate(spSize = as.numeric(str_extract(path, "^.*-(.*?)\\|.*$", group = 1))) |>   # size of the first subprob
  rowwise() %>% 
  mutate(aveWidth = mean(c_across(starts_with("width")), na.rm = T),
         points_hypercube = if_else(is.nan(aveWidth), NA, prod(1 + c_across(starts_with("width")), na.rm = T))
         #spAveCard = mean(c_across(starts_with("spCard")), na.rm = T)
         ) %>% 
  ungroup() %>% 
  mutate(
         pctSup = supported/card,
         pctExt = extreme/card,
         m = as.numeric(str_extract(path, "^.*-(.*)_.*$", group = 1))) 


## is the hypercube close to approx?
#dat %>% mutate(cst = points_hypercube/(m * 10000)^p) %>% ungroup() %>% summarise(val = mean(cst, na.rm = T))

# add relative stat for subproblems
for (s in 1:max(dat$m)) {
  dat[, str_c("spPctExt", s)] <- dat[, str_c("spExtreme", s)] / dat[, str_c("spCard", s)]
} 
dat <- dat %>% 
  rowwise() %>% 
  mutate(
    spAvePctExt = mean(c_across(starts_with("spPctExt")), na.rm = T),
    ubCard = prod(c_across(starts_with("spCard")), na.rm = T)
  ) %>% 
  ungroup()
datFull <- dat |> filter(spSize <= 300, str_detect(path, "alg1-"))
# remove results where have not solved all instances
dat <- datFull %>% 
  filter(!is.na(card)) |> 
  group_by(p, m, method, spAveCard) %>% 
  filter(n() == 5) %>% 
  ungroup()
```

`r datFull |> filter(!is.na(card)) |> nrow()`/`r nrow(datFull)` problems have been solved, i.e. `r nrow(datFull) - datFull |> filter(!is.na(card)) |> nrow()` remaining:

```{r}
datFull |> filter(is.na(card)) |> pull(path)
```

`r nrow(dat)`/`r datFull |> filter(!is.na(card)) |> nrow()` problems have 5 instances solved for each configuration. Configurations with lees that 5 solved:

```{r}
datFull |> 
  group_by(p, m, method, spAveCard) %>% 
  summarise(solved = sum(!is.na(card))) |> 
  filter(solved != 5) |> 
  print(n=100)
```


`r datFull |> filter(is.na(extreme) & !is.na(card)) |> nrow()`/`r datFull |> filter(!is.na(card)) |> nrow()` have not been classified at all:

```{r}
datFull |> filter(is.na(extreme) & !is.na(card)) |> pull(path)
```

`r datFull |> filter(is.na(unsupported) & !is.na(extreme)) |> nrow()`/`r datFull |> filter(!is.na(card)) |> nrow() - datFull |> filter(is.na(extreme) & !is.na(card)) |> nrow()` classified files have not been fully classified (only classified extreme).

  
### Problems solved for the analysis  

```{r, out.width = "100%"}
DT::datatable(dat, width = "100%", options = list(autoWidth = T, scrollX = TRUE, columnDefs = list(list(width = '250px', targets = c(1)))))
```

Note that the width of objective $i = 1, \ldots p$, $w_i = [l_i, u_i]$ should be approx. $10000m$. Check:

```{r}
dat %>% 
  group_by(m) %>% 
  summarize(across(starts_with("width"), ~ mean(.x, na.rm = TRUE), .names = "mean_{.col}"))
```



```{r, eval=FALSE}
# pts <- matrix(c(5,5,5), ncol = 3)
# ini3D(argsPlot3d = list(xlim = c(0,10), ylim = c(0,10), zlim = c(0,10)))
# plotCones3D(pts)
# plotCones3D(pts, direction = -1)
# finalize3D(argsAxes3d = list(edges = "bbox"))

## try to estimate ub on integer nd points in a hypercube by fitting pts to hyperplane

# library(eaf)
# library(OjaNP)
# estimateUbND <- function(ideal, nadir) {
#   p <- length(ideal)
#   # find lex max
#   pts <- matrix(rep(ideal, p), ncol = p, byrow = T)
#   diag(pts) <- nadir
#   # calc hyperplane containing pts
#   coeff <- hyperplane(pts)   # d^t * x + c = 0
#   # find integer pts near hyperplane
#   val <- map(1:(p-1), function(i) ideal[i]:nadir[i])
#   res <- as.matrix(expand.grid(val))
#   coeff1 <- coeff[1:(p-1)]
#   tmp <- -(coeff1 %*% t(res) + coeff[p+1])/coeff[p]
#   res <- cbind(res,tmp[1,])
#   # tst <- coeff[1:p] %*% t(tst) + coeff[p]
#   # any(tst != 0)
#   res <- res[res[,p] >= ideal[p] & res[,p] <= nadir[p], ]
#   # colnames(res) <- str_c("y", 1:(p-1))
#   # res <- res |> 
#   #   rowwise() |> 
#   #   mutate(test = -(sum(coeff[1:(p-1)] * c_across(contains("y"))) + coeff[p+1]) / coeff[p]) |> 
#   #   mutate(y = round(-(sum(coeff[1:(p-1)] * c_across(contains("y"))) + coeff[p+1]) / coeff[p])) |> 
#   #   mutate(test = sum(coeff[1:p] * c_across(contains("y"))) + coeff[p+1]) 
#   ndSet <- filter_dominated(res)
#   return(nrow(ndSet))
# }
# p <- 3
# ideal <- rep(0,p)
# nadir <- rep(10000,p)
# estimateUbND(ideal, nadir)
```



### Size of $Y_N$

What is $|Y_N|$ given the different methods of generating the set of nondominated points for the subproblems?

```{r}
dat %>% 
  group_by(method) %>% 
  summarise(mean_card = mean(card), n = n()) 
```

Does $p$ have an effect? 

```{r}
dat %>% 
  group_by(method, p) %>% 
  summarise(mean_card = mean(card), n = n()) %>% 
  arrange(p)
```

Does $m$ have an effect? 

```{r}
dat %>% 
  group_by(method, m) %>% 
  summarise(mean_card = mean(card), n = n()) %>% 
  arrange(m)
```

```{r}
dat %>% 
  ggplot(aes(x = spAveCard, y = card, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, scales = "free", labeller =  "label_both")  + 
  labs(title = "Number of nondominated points given subproblem cardinality",
       y = "Cardinality Yn",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") 
```

```{r}
dat %>% 
  ggplot(aes(x = spAveCard, y = card, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_wrap(p ~ m, scales = "free", labeller =  "label_both")  + 
  labs(title = "Number of nondominated points given subproblem cardinality",
       y = "Cardinality Yn",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") 
```

```{r}
dat %>% 
  ggplot(aes(x = spAveCard, y = card, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, scales = "free_y", labeller = "label_both")  + 
  labs(title = "Number of nondominated points given subproblem cardinality (log2) scale",
       y = "Cardinality Yn",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") +
  scale_y_continuous(trans = "log2")
```


```{r}
dat %>% 
  ggplot(aes(x = p, y = card, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(m ~ spAveCard, scales = "free_y", labeller =  "label_both")  + 
  labs(title = "Number of nondominated points given subproblem cardinality",
       y = "Cardinality Yn",
       x = "p",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") 
```

```{r}
dat %>% 
  ggplot(aes(x = m, y = card, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ spAveCard, scales = "free_y", labeller =  "label_both")  + 
  labs(title = "Number of nondominated points given subproblem cardinality",
       y = "Cardinality Yn",
       x = "m",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") 
```

```{r}
dat %>%
  ggplot(aes(x = spAveCard, y = card, color = factor(p))) +
  geom_point() +
  stat_summary(fun="mean", geom="line") +
  facet_wrap(vars(method), scales = "free", labeller = "label_both")  +
  labs(title = "Number of nondominated points given subproblem cardinality",
       y = "Cardinality Yn",
       x = "Subproblem cardinality",
       color = "p") +
  theme(legend.position = "bottom")
```

```{r}
dat %>% 
  ggplot(aes(x = spAvePctExt, y = card, color = factor(spAveCard), shape = method, group = method)) +
    stat_summary(fun="mean", geom="line", color = "gray70") +
    # stat_summary(aes(group = factor(spAveCard), color = factor(spAveCard)), fun="mean", geom="line", size = 0.05) +
    geom_point() + 
    facet_grid(p ~ m, scales = "free_y", labeller =  "label_both")  + 
    labs(title = "Number of nondominated points",
       y = "Cardinality Yn",
       x = "Ave relative number of subproblem extreme",
       color = "Subproblem cardinality") +
  theme(legend.position = "bottom") 
```

Let us try to fit the results using function $y=c_1 s^{(c_2p)} m^{c_3p}$ (different functions was tried and this gave the highest $R^2$) for each method.
```{r}
# Fit exponential regression model
library(tidymodels)
tidymodels_prefer()
lm_spec <- linear_reg()
# lm_fit <- fit(lm_spec, log(card) ~ -1 + method + method:I(p * log(spAveCard)) + method:I(p * log(m)) + I(log(spAveCard)), data = dat)
# summary(lm_fit)
# tidy(lm_fit)

regressions <- 
  dat %>%
  nest(data = c(-method)) %>% 
  mutate(
    fit = map(data, ~ lm(log(card) ~ I(p * log(spAveCard)) + I(p * log(m)), data = .x)),
    tidied = map(fit, tidy),
    glanced = map(fit, glance)
  )

regressions %>% 
  select(-data) |> 
  unnest(glanced)

# test <- function(p, spAveCard, m) {
#   regressions |> 
#     mutate(p1 = map_dbl(fit, function(x) exp(predict(x, newdata = tibble(spAveCard = spAveCard, m = m, p = p,))))) |> 
#     mutate(p2 = map_dbl(tidied, function(x) exp(x$estimate[1]) * spAveCard^(x$estimate[2] * p) * m^(x$estimate[3] * p)))
#   # cf <- coef |> filter(method == "u") |> select(-method) |> unlist()
#   # return(cf[1] * spAveCard^(cf[2] * p) * m^(cf[3] * p))
# }
# test(p = 3, spAveCard = 200, m = 3)

coef <- regressions %>% 
  select(-data) |> 
  unnest(tidied) |> 
  mutate(estimate = if_else(term == "(Intercept)", exp(estimate), estimate),
         coef = rep(c("c1", "c2", "c3"), 4)) |> 
  select(method, coef, estimate) |> 
  pivot_wider(names_from = coef, values_from = estimate)
coef

# model <- lm(log(card) ~ -1 + method + method:I(p * log(spAveCard)) + method:I(p * log(m)), data = dat)
# summary(model)
# par(mfrow=c(2,2))
# plot(model)
```



```{r}
model <- lm(log(card) ~ -1 + method + method:I(p * log(spAveCard)) + method:I(p * log(m)), data = dat)
# some plots
library(plotly)
spAveCard <- seq(100, 300, by = 10)
m <- 1:5
p <- 3
method <- factor("l", levels = unique(dat$method))
card <- t(outer(spAveCard, m, function(spAveCard, m) exp(predict(model, newdata = tibble(spAveCard = spAveCard, m = m, p = p, method = method)))))
dat1 <- expand_grid(m = m, p = p, spAveCard = spAveCard, method = method)
dat1 <- bind_cols(dat1, card = exp(predict(model, newdata = dat1)))
fig <- plot_ly() %>%
  add_surface(x = spAveCard, y = m, z = card,
    contours = list(
    x = list(show = TRUE, color = 'lightgray', width = 2),
    y = list(show = TRUE, color = 'lightgray', width = 2),
    z = list(show = TRUE, color = 'lightgray', width = 2)
  ), colorscale = "Viridis", opacity = 0.8)
  # add_trace(type = 'scatter3d',
  #           mode = 'markers',
  #           data = dat1,
  #           x = ~spAveCard, y = ~m, z = ~card,
  #           marker = list(color = 'red', size = 5, symbol = 'circle'))

method <- factor("m", levels = unique(dat$method))
card <- t(outer(spAveCard, m, function(spAveCard, m) exp(predict(model, newdata = tibble(spAveCard = spAveCard, m = m, p = p, method = method)))))
dat1 <- expand_grid(m = m, p = p, spAveCard = spAveCard, method = method)
dat1 <- bind_cols(dat1, card = exp(predict(model, newdata = dat1)))
fig <- fig |> add_surface(x = spAveCard, y = m, z = card,
  contours = list(
    x = list(show = TRUE, color = 'lightgray', width = 2),
    y = list(show = TRUE, color = 'lightgray', width = 2),
    z = list(show = TRUE, color = 'lightgray', width = 2)
  ), colorscale = 'Rainbow', opacity = 1)

method <- factor("ul", levels = unique(dat$method))
card <- t(outer(spAveCard, m, function(spAveCard, m) exp(predict(model, newdata = tibble(spAveCard = spAveCard, m = m, p = p, method = method)))))
dat1 <- expand_grid(m = m, p = p, spAveCard = spAveCard, method = method)
dat1 <- bind_cols(dat1, card = exp(predict(model, newdata = dat1)))
fig <- fig |> add_surface(x = spAveCard, y = m, z = card,
  contours = list(
    x = list(show = TRUE, color = 'lightgray', width = 2),
    y = list(show = TRUE, color = 'lightgray', width = 2),
    z = list(show = TRUE, color = 'lightgray', width = 2)
  ), colorscale = 'heat', opacity = 1)

method <- factor("u", levels = unique(dat$method))
card <- t(outer(spAveCard, m, function(spAveCard, m) exp(predict(model, newdata = tibble(spAveCard = spAveCard, m = m, p = p, method = method)))))
dat1 <- expand_grid(m = m, p = p, spAveCard = spAveCard, method = method)
dat1 <- bind_cols(dat1, card = exp(predict(model, newdata = dat1)))
fig <- fig |> add_surface(x = spAveCard, y = m, z = card,
  contours = list(
    x = list(show = TRUE, color = 'lightgray', width = 2),
    y = list(show = TRUE, color = 'lightgray', width = 2),
    z = list(show = TRUE, color = 'lightgray', width = 2)
  ), colorscale = 'Cividis', opacity = 1) |> 
  layout(
    title = list(
      x = 0.5, y = 0.9, z = 0.9, xanchor = "center",
      text = "Yn for p = 3 and method u, m and l (higest Yn)"  # Title text
    ),
    scene = list(
    xaxis = list(range = range(spAveCard)),  # Set x-axis limits
    yaxis = list(range = range(m),  # Set y-axis limits
    zaxis = list(range = range(card))     # Set z-axis limits
  ))) 

fig
```



```{r}
spAveCard <- seq(100, 300, by = 10)
m <- 1:5
p <- 4
method <- factor("m", levels = unique(dat$method))
card <- t(outer(spAveCard, m, function(spAveCard, m) exp(predict(model, newdata = tibble(spAveCard = spAveCard, m = m, p = p, method = method)))))
dat1 <- expand_grid(m = m, p = p, spAveCard = spAveCard, method = method)
dat1 <- bind_cols(dat1, card = exp(predict(model, newdata = dat1)))
fig <- plot_ly() %>%
  add_surface(x = spAveCard, y = m, z = card,
    contours = list(
    x = list(show = TRUE, color = 'lightgray', width = 2),
    y = list(show = TRUE, color = 'lightgray', width = 2),
    z = list(show = TRUE, color = 'lightgray', width = 2)
  ), opacity = 0.8)

p <- 3
card <- t(outer(spAveCard, m, function(spAveCard, m) exp(predict(model, newdata = tibble(spAveCard = spAveCard, m = m, p = p, method = method)))))
dat1 <- expand_grid(m = m, p = p, spAveCard = spAveCard, method = method)
dat1 <- bind_cols(dat1, card = exp(predict(model, newdata = dat1)))
fig <- fig |> add_surface(x = spAveCard, y = m, z = card,
  contours = list(
    x = list(show = TRUE, color = 'lightgray', width = 2),
    y = list(show = TRUE, color = 'lightgray', width = 2),
    z = list(show = TRUE, color = 'lightgray', width = 2)
  ), opacity = 0.8) 

p <- 2
card <- t(outer(spAveCard, m, function(spAveCard, m) exp(predict(model, newdata = tibble(spAveCard = spAveCard, m = m, p = p, method = method)))))
dat1 <- expand_grid(m = m, p = p, spAveCard = spAveCard, method = method)
dat1 <- bind_cols(dat1, card = exp(predict(model, newdata = dat1)))
fig |> add_surface(x = spAveCard, y = m, z = card,
  contours = list(
    x = list(show = TRUE, color = 'lightgray', width = 2),
    y = list(show = TRUE, color = 'lightgray', width = 2),
    z = list(show = TRUE, color = 'lightgray', width = 2)
  ), opacity = 0.8) |> 
  layout(
    title = list(
      x = 0.5, y = 0.9, z = 0.9, xanchor = "center",
      text = "Yn for p = 2-4 and method m"  # Title text
    ),
    scene = list(
    xaxis = list(range = range(spAveCard)),  # Set x-axis limits
    yaxis = list(range = range(m),  # Set y-axis limits
    zaxis = list(range = range(card))     # Set z-axis limits
  ))) 
```




### Relative size of $Y_N$

```{r}
dat %>% 
  ggplot(aes(x = spAveCard, y = card / ubCard, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, labeller = "label_both")  + 
  labs(title = "Relative size of nondominated points compared to total enumeration given subproblem cardinality",
       y = "Cardinality Yn",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") 
```

```{r, eval=FALSE}
dat %>% 
  ggplot(aes(x = spAveCard, y = card / ubCard, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, labeller = "label_both")  + 
  labs(title = "Relative size of nondominated points compared to total enumeration given subproblem cardinality (log2 scale)",
       y = "Cardinality Yn",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") +
  scale_y_continuous(trans = "log2")
```

```{r}
dat %>% 
  ggplot(aes(y = card / ubCard, x = method, color = method)) +
  geom_boxplot() +
  facet_grid(p ~ m, labeller = label_both) +
  labs(title = "Relative size of nondominated points compared to total enumeration",
       y = "%",
       x = "Method",
       color = "Method(s) used to generate subproblems") +
  theme(legend.position = "bottom")
```


```{r, eval=FALSE}
dat %>% 
  ggplot(aes(y = card / ubCard, x = method, color = method)) +
  geom_boxplot() +
  facet_grid(m ~ p, labeller = label_both) +
  labs(title = "Relative size of nondominated points compared to total enumeration (log2 scale)",
       y = "%",
       x = "Method",
       color = "Method(s) used to generate subproblems") +
  theme(legend.position = "bottom") +
  scale_y_continuous(trans = "log2")
```


<!-- ```{r} -->
<!-- dat %>%  -->
<!--   ggplot(aes(y = card / points_hypercube, x = method, color = method)) + -->
<!--   geom_boxplot() + -->
<!--   facet_grid(m ~ p, margins = T, labeller = label_both) + -->
<!--   labs(title = "Relative size of nondominated points compared to total enumeration", -->
<!--        y = "%", -->
<!--        x = "Method", -->
<!--        color = "Method(s) used to generate subproblems") + -->
<!--   theme(legend.position = "bottom") -->
<!-- ``` -->


### Nondominated points classification

We classify the nondominated points into, extreme, supported non-extreme and unsupported.

```{r}
dat %>% 
  ggplot(aes(x = spAveCard, y = pctExt, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, margins = T, labeller = label_both)  + 
  labs(title = "Relativ number of extreme points given subproblem cardinality",
       y = "%",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") 
```

```{r}
dat %>% 
  ggplot(aes(x = p, y = pctExt, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(spAveCard ~ m, margins = T, labeller = label_both)  + 
  labs(title = "Relativ number of extreme points given subproblem cardinality",
       y = "%",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") 
```

```{r}
dat %>% 
  ggplot(aes(x = m, y = pctExt, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ spAveCard, margins = T, labeller = label_both)  + 
  labs(title = "Relativ number of extreme points given subproblem cardinality",
       y = "%",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") 
```

```{r}
dat |> 
  summarise(minPctEx = min(pctExt, na.rm = T), avePctExt = mean(pctExt, na.rm = T), maxPctEx = max(pctExt, na.rm = T)) 

dat |> 
  group_by(method) |> 
  summarise(minPctEx = min(pctExt, na.rm = T), avePctExt = mean(pctExt, na.rm = T), maxPctEx = max(pctExt, na.rm = T)) |> 
  arrange(desc(avePctExt))
```


```{r}
dat %>% 
  ggplot(aes(x = spAveCard, y = pctSup, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, margins = T, labeller = label_both)  + 
  labs(title = "Relativ number of supported points given subproblem cardinality",
       y = "%",
       x = "Subproblem cardinality",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom") 
```

```{r}
dat %>% 
  ggplot(aes(x = method, y = pctExt, color = method)) +
  geom_boxplot() + 
  facet_grid(p ~ m, margins = T, labeller = label_both) +
  labs(title = "Relativ number of extreme points given method",
       y = "%",
       x = "Method",
       color = "Methods used to generate subproblems") +
  theme(legend.position = "bottom")
```

```{r}
dat %>% 
  ggplot(aes(x = spAvePctExt, y = pctExt, color = method)) +
  geom_point() + 
  facet_grid(p ~ m, margins = T, labeller = label_both) +
  labs(title = "Relativ number of extreme points",
       y = "% extreme master",
       x = "Avg. % extreme subproblems",
       color = "Method(s) used to generate subproblems") +
  theme(legend.position = "bottom")
```


```{r}
dat %>% 
  ggplot(aes(x = spAvePctExt, y = pctExt, color = factor(spAveCard), shape = method)) +
  geom_point() + 
  facet_grid(p ~ m, margins = T, labeller = label_both) +
  labs(title = "Relativ number of extreme points",
       y = "% extreme master",
       x = "Avg. % extreme subproblems",
       color = "Subproblem cardinality") +
  theme(legend.position = "bottom")
```


```{r}
dat %>% 
  ggplot(aes(y = pctExt/spAvePctExt, x = spAveCard, color = method)) +
  geom_point() + 
  stat_summary(fun="mean", geom="line") +
  facet_grid(p ~ m, margins = T, labeller = label_both) +
  labs(title = "Relativ: % extreme master / % extreme subproblems",
       y = "% extreme master / % extreme subproblems",
       x = "Subproblem cardinality",
       color = "Method(s) used to generate subproblems") +
  theme(legend.position = "bottom")
```



## Plots used in the paper

```{r, eval=FALSE}

# Number of nondominated points given subproblem cardinality"
pt1 <- dat %>% 
  mutate(str_lbl = str_c("$p=", p, "$", " $|S|=", m, "$")) |> 
  mutate(method = case_when(
    method == "l" ~ "\\mL",
    method == "m" ~ "\\mM",
    method == "u" ~ "\\mU",
    method == "ul" ~ "\\mL{}\\mU",
    .default = "error"
  )) |> 
  ggplot(aes(x = spAveCard, y = card, shape = method, linetype = method)) +
  geom_point(alpha = 0.3) + 
  stat_summary(fun="mean", geom="line") +
  facet_wrap(vars(str_lbl), scales = "free_y")  + 
  labs(x = "$|\\Yn^s|$", y = "$|\\Yn|$", shape = NULL, linetype = NULL) +
  scale_linetype_manual(values = c("solid", "dashed", "dotted", "dotdash")) +
  scale_y_continuous(labels = scales::number_format(big.mark = ",", prefix = "\\ ")) +
  guides(linetype = guide_legend(override.aes = list(size = 0, keywidth = 16))) +
  theme_bw() +
  theme(legend.position = "bottom", legend.key.width = unit(2, "cm"))
pt1

tikz("~/Library/CloudStorage/Dropbox/Apps/Overleaf/paper1/figures/yn-spcard.tex", width = 10, height = 10)
pt1
dev.off()


## Number of extreme
dat |> 
  filter(!is.na(extreme)) |> 
  mutate(e = extreme/card) |> 
  # group_by(method, p) |> 
  group_by(method) |> 
  summarise(n = n(), eMax = max(e), eMin = min(e), e = mean(e)) |> 
  arrange(desc(e)) 


# pt1 <- dat |> 
#   ggplot(aes(x = card, y = extreme/card, shape = method, linetype = method)) +
#   stat_summary(fun="mean", geom="line") + 
#   geom_point(size = 0.75) + 
#   facet_grid(cols = vars(p), scales = "free") +
#   labs(x = "$\\Yn^s$", y = NULL) +
#   theme_bw() +
#   theme(legend.position = "none")
# pt1
# 
# pt2 <- dat |> 
#   ggplot(aes(x = p, y = extreme/card, shape = method, linetype = method)) +
#   geom_point(size = 0.75) + 
#   stat_summary(fun="mean", geom="line") + 
#   facet_grid(cols = vars(card)) + 
#   labs(x = "$p$", y = NULL) +
#   theme_bw() +
#   theme(legend.position = "none")
# pt2
# 
# tikz("se-card.tex", width = 8, height = 3)
# pt1
# dev.off()
# tikz("se-p.tex", width = 8, height = 3)
# pt2
# dev.off()
```
