---
title: "Generation of instance files for subproblems"
output: html_document
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Subproblem caristica

\begin{itemize}
    \item Test $p=2,\ldots, 5$. [4 options]
    \item Width is the same for all objectives (\cite{Kerberenes2022phd} use $w_i=1000$ in Chap. 2) or width is the same for approx. half of the objectives (\eg 1000) and 1/4 for the remaining (\ie 250). [2 options]
    \item Generate on a sphere: either lower/west part (many supported) or upper/east part (many unsupported). [2 options]
    \item Generate $10, 50, 100, 150, 200$ points and find the nd points among them for each subproblem. [5 options]
    \item Classify each nd point (supported, extreme, unsupported)
    \item Keep statistics: number of points, supported, extreme, unsupported, ratio = supported/unsupported, (range) width $w_i$ = max-min, min and max value for each objective $i = 1,\ldots p$.
    \item 5 instances for each config. [5 options]
\end{itemize}

```{r}
#install.packages("gMOIP")
library(gMOIP)
```

```{r}
ini3D()
cent <- c(100,100,100)
r <- 75
planeC <- c(cent+r/3)
planeC <- c(planeC, -sum(planeC^2))
pts <- genSample(3, 100,
  argsSphere = list(center = cent, radius = r, below = FALSE, plane = planeC, factor = 6))
Rfast::colMinsMaxs(as.matrix(pts))
rgl::spheres3d(cent, radius=r, color = "grey100", alpha=0.1)
plotPoints3D(pts)
rgl::planes3d(planeC[1],planeC[2],planeC[3],planeC[4], alpha = 0.5, col = "red")
finalize3D()
```

